### 2017-10-12 paper/sections/4_methodology.tex
paper: add template
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Methodology: Global Namespace with Per-Subtree Consistency/Durability}
\label{sec:methodology-decoupled-namespaces}

\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\caption{Applications decouple the namespace, write updates to a local journal,
and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
\end{figure}

\begin{table}
\begin{tabular}{ r | l }
  Mechanism         & Description \\\hline
  RPCs              & round trip remote procedure calls \\
  Stream            & stream journal into object store \\
  Append Client Journal & events appended to in-memory journal \\
  Volatile Apply    & apply to metadata store in obj store \\
  Nonvolatile Apply & apply to metadata store in memory \\
  Local Persist     & journal saved to client's disk \\
  Global Persist    & journal saved in object store \\
\end{tabular}
\caption{The CudeleFS mechanisms are composed together to form consistency and
durability semantics.\label{table:mechanisms}} 
\end{table}

In this section we describe CudeleFS, our file system that lets users
assign consistency and durability semantics to subtrees in the global
namespace. A \textbf{mechanism} is an abstraction and basic building block for
constructing consistency and durability guarantees. CudeleFS exposes these
mechanisms and the user composes them together to construct
\textbf{policies}. These policies are assigned to subtrees and they dictate how
the file system should handle operations within that subtree.  Below, we
describe the mechanisms, the policies, and the API for assigning policies to
subtrees.

\subsection{The CudeleFS Mechanisms}
\label{sec:the-cudelesfs-mechanisms}

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in CudeleFS and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Decoupled clients use the ``Append Client Journal"
mechanism to append metadata updates to a local, in-memory journal. Once the
job is complete, the system calls CudeleFS mechanisms to achieve the desired
consistency/durability semantics.  CudeleFS provides a library for clients to
link into and all operations are performed by the client.

\subsubsection{Mechanisms Used for Consistency} ``RPCs" send remote procedure
calls for every metadata operation from the client to the metadata server,
assuming the request cannot be satisfied by the inode cache. This mechanism is
part of the default CephFS implementation and is the strongest form of
consistency because clients see metadata updates right away.  ``Nonvolatile
Apply" replays the client's journal onto the metadata cluster's metadata store.
The client's in-memory journal is written into the object store and the
metadata servers are restarted. When the metadata servers re-initialize, they
notice new journal updates in the object store and replay the events onto their
in-memory metadata stores.  ``Volatile Apply" takes the client's in-memory journal on the
client and applies the updates directly to the in-memory metadata store maintained
by the metadata servers. We say volatile because -- in exchange for peak
performance -- CudeleFS makes no consistency or durability guarantees while
``Volatile Apply" is executing.  If a concurrent update from a client occurs
there is no rule for resolving conflicts and if the client or metadata server
crashes there may be no way to recover.

% difference between apply and volatile apply
The biggest difference between ``Volatile Apply" and ``Nonvolatile Apply" is
the medium they use to communicate. ``Volatile Apply" applies updates directly
to the metadata servers' metadata store while ``Nonvolatile Apply" uses the
object store to communicate the journal of updates from the client to the
metadata servers.  ``Nonvolatile Apply" is safer but has a large performance
overhead because objects in the metadata store need to be read from and written
back to the object store.

%The metadata store and journal
%are different ways of representing the namespace.  CudeleFS presents 6
%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
%Persist. ``RPCs" does round trip remote procedure calls to establish
%consistency; it is the default implementation for complying with POSIX in
%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
%into the object store. ``Append Client Journal" allows clients to append metadata events to an
%in-memory journal. ``Volatile apply" 
% ``Local Persist" takes the in-memory journal and writes it to the
%client's disk. ``Global Persist" saves the journal as a an object in the object
%store from the client.  Next, we discuss how these mechanisms can be composed
%to get different consistency and durability semantics.

\subsubsection{Mechanisms Used for Durability} ``Stream" is one of the
mechanisms used by default in CephFS.  Using existing configuration settings in
Ceph we can turn ``Stream" on and off.  If it is off, then the metadata servers
will not save journals in the object store. For ``Local Persist", clients write
serialized log events to a file on local disk and for ``Global Persist",
clients push the journal into the objects store. The overheads for both ``Local
Persist" and ``Global Persist" is the write bandwidth of the local disk and
object store, respectively.  These persist mechanisms are part of the library that
links into the client.

\subsection{Defining Policies in CudeleFS}
\label{sec:setting-policies-with-cudele}

\begin{table}[t]
\begin{center}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
  none                       & append client journal            & append client journal          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & append client journal            & append client journal          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & append client journal            & append client journal          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}

\caption{Users can explore the consistency (C) and
durability (D) spectrums by composing CudeleFS mechanisms. 
\label{table:spectrum}}
\end{center}
\end{table}

% describe table
The spectrum of consistency and durability guarantees that users can
construct is shown in Table~\ref{table:spectrum}. The columns are the different
consistency semantics and the rows cover the spectrum of durability guarantees.
For consistency: ``invisible" means the system does not handle merging updates
into a global namespace and it is assumed that middleware or the application
manages consistency lazily; ``weak" merges updates at some time in the
future ({\it e.g.}, when the system has time, when the number of updates reaches a
certain threshold, when the client is done writing, etc.); and updates in
``strong" consistency are seen immediately by all clients. For durability,
``none" means that updates are volatile and will be lost on a failure. Stronger
guarantees are made with ``local", which means updates will be retained if the
client node recovers, and ``global", where all updates are always recoverable.

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability ~\footnote{This is the normal case.
IndexFS also has bulk merge which would transition the system into ``weak
consistency"}; DeltaFS uses ``invisible" consistency and ``local" durability
and BatchFS uses ``weak" consistency and ``local" durability. These systems
have other features that could push them into different semantics but we assign
labels here based on the points emphasized in the papers.  To compose the
mechanisms users inject which mechanisms to run and which to use in parallel
using a domain specific language.  Although we can achieve all permutations of
the different guarantees in Table~\ref{table:spectrum}, not all of them make
sense. For example, it makes little sense to do \texttt{append client journal+RPCs} since
both mechanisms do the same thing or \texttt{stream+save} since ``global"
durability is stronger and has more overhead than ``local" durability.

% talk of eventual consistency
The consistency and durability properties in Table~\ref{table:spectrum} are not
guaranteed until all mechanisms in the cell are complete. The compositions
should be considered atomic and there are no guarantees while transitioning
between policies. For example, updates are not deemed to have ``global" durability until
they are safely saved in the object store. If a failure occurs during ``global
persist" or if we inject a new policy that changes a subtree from ``local
persist" to ``global persist", CudeleFS makes no guarantee until the mechanisms are
complete.

\subsection{CudeleFS Namespace API}
\label{sec:cudelefs-namespace-api}

% the interface
Users control consistency and durability for subtrees by contacting the monitor
daemon with a directory path and presenting a policies configuration file.
For example, (msevilla/mydir, policies.yml) would decouple the path
``msevilla/mydir" and would apply the policies in ``policies.yml".  The
policies file supports the following parameters (default values are in parenthesis):

\begin{itemize}

  \item \texttt{consistency}: which consistency model to use (\texttt{RPCs})

  \item \texttt{durability}: which durability model to use (\texttt{stream})

  \item \texttt{allocated\_inodes}: the number of inodes to provision to the
  decoupled namespace (100)

  \item \texttt{interfere\_policy}: how to handle a request from another
  client targeted at this subtree (\texttt{allow}))

\end{itemize}

The ``Consistency" and ``Durability" parameters are set to the CudeleFS
mechanisms; they can be composed (\(+\)) or run in parallel (\(|\|\)).
``Allocated Inodes" is a way for the application to specify how many files it
intends to create. It is a contract so that the file system can provision
enough resources for the incumbent merge and so it can give valid inodes to
other clients. The inodes can be used anywhere within the decoupled namespace
({i.e.} at any depth in the subtree).

% description
``Interfere Policy" has two settings: \texttt{block} and \texttt{allow}.
For \texttt{block}, any requests to this part of the namespace returns with
``Device is busy", which will spare the metadata server from wasting resources
for updates that may get overwritten. If the application does not mind losing
updates, for example it wants approximations for results that take too long to
compute, it can select \texttt{allow}. In this case, metadata from the interfering client will be
written and the computation from the decoupled namespace will take priority at
merge time because the results are more accurate.

% examples
Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. To implement DeltaFS on CudeleFS, the user
would use the configuration:

\begin{listing}[h]
\begin{minted}[xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "append_client_journal"
    "durability": "local_persist"
}
\end{minted}
\end{listing}

BatchFS merges updates into the global namespace after the job is complete.
This can be achieved with by composing two mechanisms:

\begin{listing}[h]
\begin{minted}[xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
  "allocated_inodes": "100000"
  "interfere_policy": "block"
  "consistency": "append_client_journal+volatile_apply"
  "durability": "local_persist"
}
\end{minted}
\end{listing}