### 2017-10-23 paper/sections/3_posix_overheads.tex
paper: tone down HPCness and simplify captions
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{POSIX IO Overheads}
\label{sec:posix-overheads}

\begin{figure}[tb]
\centering
\includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
\caption{Create-heavy workloads (such as \texttt{untar}) incur the highest disk, network, and
CPU utilization because of the consistency and durability demands of
CephFS.}\label{fig:overhead-creates}
\end{figure}

In our examination of the overheads of POSIX IO we benchmark and analyze
CephFS, the file system that uses Ceph's object store ({\it i.e.} RADOS) to
store its data and metadata. To show how the file system behaves under high
metadata load we use a create-heavy workload.  During this process we
discovered, based on the analysis and breakdown of costs, that durability and
consistency have high overhead but we urge the reader to keep in mind that this
file system is an implementation of one set of design decisions and our goal
here is to highlight the effect that those decisions have on performance.

Figure~\ref{fig:overhead-creates} shows the resource utilization of compiling
the Linux kernel.  The \texttt{untar} phase, which is characterized by many
creates, has the highest resource usage (combined CPU, network, and disk)
because of the number of RPCs needed for consistency and durability. The RPCs
are also serialized because the metadata server is single threaded; although
na\"{\i}ve, this design is common because of the complexity of multi-threaded
metadata servers~\cite{konstantinos:pdsw2014-lustre-metadata}.  Traditional
file system techniques for improving performance, such as caching inodes, do
not help for create-heavy workloads.  In this section, we quantify the costs of
consistency and durability in CephFS.  At the end of each subsection we compare
the approach to ``decoupled namespaces", the technique in related work that
detaches subtrees from the global namespace to relax consistency/durability
guarantees.

\subsection{Durability}
\label{sec:durability}

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
      \caption{Runtime normalized to no journal} \label{fig:overhead-a}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.2\linewidth]{graphs/behavior-interfere.png}
      \caption{Interference forces \texttt{lookup()}s}
      \label{fig:overhead-b}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere-scale.png}
      \caption{Interference hurts variability}
      \label{fig:overhead-c}
  \end{subfigure}
  \caption{The overhead of durability and strong consistency in CephFS.
  (a) shows the effect of different journal segment sizes, which are streamed
  into the object store for fault tolerance. (b) and (c) show that when a second
  client ``interferes", capabilities are revoked and metadata servers do more
  work.  \label{fig:overhead}}
\end{figure*}

% what is durability
While durability is not specified by POSIX IO, users expect that files they
create or modify survive failures.  We define three types of durability:
global, local, and none.  Global durability means that the client or server can
fail at any time and metadata will not be lost because it is ``safe" ({\it
i.e.} striped or replicated across a cluster). Local durability means that
metadata can be lost if the client or server stays down after a failure. None
means that metadata is volatile and that the system provides no guarantees when
clients or servers fail.  None is different than local durability because
regardless of the type failure, metadata will be lost when components die in a
None configuration.

% - sequential IO, trim redundant operations
\textbf{CephFS Design}: A journal of metadata updates that streams into the
resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to be large
(on the order of MBs) which ensures (1) sequential writes into the object store
and (2) the ability for daemons to trim redundant or irrelevant journal
entries.  The journal is striped over objects where multiple journal updates
can reside on the same object. There are two tunables for controlling the
journal: the segment size and the number of segments that can be written in
parallel. Unless the journal saturates memory or CPU resources, larger values
for these tunables results in better performance.

% purpose of the journal
In addition to the metadata journal, CephFS also represents metadata in RADOS
as a metadata store, where directories and their file inodes are stored as
objects.  The metadata server applies the updates in the journal to the
metadata store when the journal reaches a certain size. The metadata store is
optimized for recovery ({\it i.e.} reading) while the metadata journal is
write-optimized.

%\begin{figure}[tb] \centering
%\includegraphics[width=1\linewidth]{./figures/journal.png} 
%\caption{CephFS has two views of the file system namespace: a journal of
%metadata updates and a metadata store. For fault tolerance, they are stored in
%the object store as segments and fragments, respectively.
%\label{fig:journal}}
%\end{figure}

% Effects on performance
Figure~\ref{fig:overhead-a} shows the
effect of journaling of different journal segment sizes; the larger the segment
size the bigger that the writes into the object store are. The trade-off for
better performance is memory consumption because larger segments take up
more space for buffering. When journaling is on, 
the metadata server periodically stops serving requests to flush ({\it i.e.}
apply journal updates) to the metadata store.  The journal overhead is
sufficient enough to slow down metadata throughput but not so much as to
overwhelm the bandwidth of the object store. We measured our peak bandwidth to
be 100MB/s, which is the speed of our network link.

\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS, to the
best of our knowledge, when a client or server fails there is no recovery
scheme. For BatchFS, if a client fails when it is writing to the local
log-structured merged tree (implemented as an
SSTable~\cite{ren:atc2013-tablefs}) then those batched metadata operations are
lost. For DeltaFS, if the client fails then, on restart, the computation does
the work again -- since the snapshots of the namespace are never globally
consistent and there is no ground truth.  On the server side, BatchFS and
DeltaFS use IndexFS. IndexFS writes metadata to SSTables, which initially
reside in memory but are later flushed to the underlying distributed file
system.

\subsection{Strong Consistency}
\label{sec:strong-consistency}

Access to metadata in a POSIX IO-compliant file system is strongly consistent, so
reads and writes to the same inode or directory are globally ordered.  The
synchronization and serialization machinery needed to ensure that all clients
see the same state has high overhead.

\textbf{CephFS Design}: Capabilities keep metadata strongly
consistent. To reduce the number of RPCs needed for consistency, clients can
obtain capabilities for reading, creating and updating inodes, as well as caching reads,
buffering writes, changing the file size, and performing lazy IO.

% inode cache - reduces RPCs (lookups for create, readdirs for stats)
To keep track of the read caching and write buffering capabilities, the clients
and metadata servers agree on the state of each inode using an inode cache.  If
a client has the directory inode cached it can do metadata writes ({\it e.g.},
create) with a single RPC. If the client is not caching the directory inode
then it must do an extra RPC to determine if the file exists.  Unless the
client immediately reads all the inodes in the cache ({\it i.e.} \texttt{ls
-alR}), the inode cache is less useful for create-heavy workloads.

% benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:overhead-b}. The colors show the behavior of the client for two
different runs.  If only one client is creating files in a directory
(``isolated" curve on \(y1\) axis) then that client can lookup the existence of
new files locally before issuing a create request to the metadata server. If
another client starts creating files in the same directory (``interfere" curve
on \(y1\) axis) then the directory inode transitions out of read caching and
the first client must send \texttt{lookup()}s to the metadata server
(``interfere" curve on \(y2\) axis). These extra requests increase the
throughput of the ``interfere" curve because the metadata server can handle the
extra load but performance suffers.  This degradation is shown in
Figure~\ref{fig:overhead-c}, where we scale the number of clients and show the
slowdown of the slowest client.  The results are normalized to a single
isolated client without a metadata server journal.  For the ``interfere" bars,
each client creates files in private directories and at 30 seconds we launch
another process that creates files in those directories. 18 clients is the
maximum number the metadata server could handle for this metadata-intensive
workload; at higher client load, the metadata server started complaining about
laggy and unresponsive requests.

% TODO: what is the cost of trimming the cache?
% TODO: does CephFS still cache inodes when I turn off caching? Why is still keeping inodes in memory? Gah!

\textbf{Comparison to decoupled namespaces}: Decoupled namespaces
merge batches of metadata operations into the global namespaces when the job
completes.  In BatchFS the merge is delayed by the application using an API to
switch between asynchronous to synchronous mode. The merge itself is explicitly
managed by the application but future work looks at more automated
methodologies. In DeltaFS snapshots of the metadata subtrees stays on the client
machines; there is no ground truth and consistent namespaces are constructed
and resolved at application read time or when a 3rd party system ({\it e.g.},
middleware, scheduler, etc.) needs a view of the metadata. As a result, all the
overheads of maintaining consistency that we showed above are delayed until the
merge phase.