### 2017-12-10 paper/cover_letter.tex
paper: add cover letter skeleton (#25)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% Template: http://www.acm.org/publications/proceedings-template
\documentclass[onecolumn,conference]{IEEEtran}
\usepackage{booktabs} % For formal tables
\usepackage{minted}
\usepackage{graphicx}
\usepackage{arydshln}
\usepackage{subcaption}
\usepackage{times}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\begin{document}
\title{Cover Letter}
\maketitle

We thank the reviewers for their hard work and genuinely helpful suggestions.
In addition to this cover letter, we have posted a PDF online explicitly
showing additions in \textcolor{red}{\textbf{red}} and deletions in
\textcolor{blue}{\textbf{blue}}. Three issues were common across reviewer
feedback:

%Experimental Setup (R1, R2, R3)
%- code R1
%- servers, network, storage R1, R3
\section{No Experimental Setup (Reviewers 1, 2, and 3)}

We apologize for the omission of experimental setup and source code details.
For the experimental setup, we have added details about the cluster we ran on
in Table III as well as a description in the text (addresses Reviewers 1 and
3). For the source code, we have added code snippets to the Evaluation to
concretize what was actually done (addresses Reviewer 1).

Additionally, we have added paper artifacts online.  For the experimental
setup, we have made the infrastructure code available and added links after
each figure to show exactly how experiments are run.  This infrastructure code
contains scripts to deploy the system, run experiments, and gather results.
This process follows the Popper
Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
aims to make research reproducible.  We also make the source code available and
linked in the paper.

%Mixing contributions and future work (R1)
%Statements in Introduction do not align with Evaluation (R1, R2, R3)
%- quantify performance speedups R3
%Evaluation Structure (R2)
%- remove major takeaways and cross refercnes
%Will not address:
%- cost of dynamically changing consistency/durability not presented R1
\section{Structure and Layout of Evaluation (All Reviewers)}

One pervasive issue across reviews is that the results and contributions we
cite in the introduction are not validated, explained, or even mentioned in the
Evaluation.  To fix this, we separate out future work to make the contributions
of this paper more explicit (addresses Reviewer 1) and we connect all numbers
we cite in the introduction to the evaluation (addresses all reviewers).  To
further clarify the Evaluation, we:

\begin{itemize}

  \item added Table IV which includes raw baseline, Cudele, and hardware limit
numbers. This adds raw numbers and helps us make comparisons to hardware
speeds. Ultimately, we hope this adds insight into the results (addresses
Reviewers 2 nad 3).

  \item re-organized the section to make experiments more independent
and removed cross-references (addresses Reviewer 2). Experiments are now
self-contained so the reader can see the effects of different API
configurations (addresses Reviewer 2) and we do a better job of explaining how
results are derived (addresses Reviewer 3).

  \item removed the ``major takeaways". We deleted the headings but left the
conclusions we make as transitions (Reviewer 2). We also add insights into the
results by analyzing the raw numbers we observe.

\end{itemize}

%Use Cases R2, R4
%- Spark discussion R2
%- more emphasis on parallel and distributed computing themes R4
\section{Synthetic and/or Irrelevant Use Cases (Reviewers 2 and 4)}

We add a section to the introduction that scopes the importance of storage for
large scale runtimes and workflows. Parallel and distributed jobs (whether that
be MPI-based or more structured runtimes like Hadoop) need parallel and
distributed storage systems that keep up with the computation. Furthermore, we
highlight the parallel and distributed computing themes (addresses Reviewer 4)
and cast light on them from the context of our prototype. We also add the Spark
discussion proposed by Reviewer 2 to the Evaluation. The goal is to show how a
system such as Cudele would benefit more general runtimes. We also move the
discussion about the potential for HDFS and CephFS subtrees in the same
namespace and the cost of dynamically changing consistency and durability
semantics on a subtree.

\section*{Cosmetic Fixes}
%Terminology
%- users vs. clients R2

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,paper}
\end{document}