### 2017-12-13 paper/cover_letter.tex
paper: fix graphs and add baselines (#33, #34)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% Template: http://www.acm.org/publications/proceedings-template
\documentclass[onecolumn,conference]{IEEEtran}
\usepackage{booktabs} % For formal tables
\usepackage{minted}
\usepackage{graphicx}
\usepackage{arydshln}
\usepackage{subcaption}
\usepackage{times}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\begin{document}
\title{Cover Letter}
\maketitle

We thank the reviewers for their hard work and genuinely helpful suggestions.
In addition to this cover letter, we have posted a PDF online explicitly
showing additions in \textcolor{red}{\textbf{red}} and deletions in
\textcolor{blue}{\textbf{blue}}. Three issues were common across reviewer
feedback; Reviewer 1 and 3's comments are addressed in issues I and II,
Reviewer 2's comments are addressed in issues I and III, and Reviewer 4's
comments are addressed in issues II and III.

%Experimental Setup (R1, R2, R3)
%- code R1
%- servers, network, storage R1, R3
\section*{Issue I: No Experimental Setup (Reviewers 1, 2, and 3)}

We apologize for the omission of experimental setup and source code details. To
provide a more comprehensive view of our experiments, we added text to the
paper and made paper artifacts available online. For the experimental setup, we
describe the cluster setup (hardware, software, etc.) in Section~{\S}V.  We
have also made the infrastructure code available and added links after each
figure to show exactly how experiments are run.  This infrastructure code
contains scripts to deploy the system, run experiments, and gather results.
This process follows the Popper
Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
aims to make research reproducible.  For the source code, we have added code
snippets to the evaluation to concretize what was actually done and made the
source code available online. A link is provided in a footnote in
Section~{\S}V.

%Mixing contributions and future work (R1)
%Statements in Introduction do not align with Evaluation (R1, R2, R3)
%- quantify performance speedups R3
%Evaluation Structure (R2)
%- remove major takeaways and cross refercnes
%Will not address:
%- cost of dynamically changing consistency/durability not presented R1
\section*{Issue II: Structure and Layout of Evaluation (All Reviewers)}

All reviewers note that the results and contributions we cite in the
introduction are not validated, explained, or even mentioned in the evaluation.
One confusing component of this issue is that we mixed future work with the
contributions of this paper; to make the contributions more explicit we remove
future work from the introduction and add it to the future work section in
Section~{\S}VI. We also connect all numbers we cite in the introduction to the
evaluation. To further clarify the evaluation, we:

\begin{itemize}

  \item added raw numbers (as annotations) to the baselines in each figure.
This helps the reader calculate the raw numbers instead of just relying on
overhead and speedup numbers.

  \item re-organized the section to make experiments more independent and
removed cross-references. Experiments are now self-contained so the reader can
see the effects of different API configurations and we do a better job of
explaining how results are derived.

  \item removed the ``major takeaways". We deleted the headings but left the
conclusions we make as transitions. We also add insights into the results by
analyzing the raw numbers we observe in comparison to hardware capabilities.

\end{itemize}

%Use Cases R2, R4
%- Spark discussion R2
%- more emphasis on parallel and distributed computing themes R4
\section*{Issue III: Synthetic and/or Irrelevant Use Cases (Reviewers 2 and 4)}

We add a section to the introduction that scopes the importance of storage for
large scale runtimes and workflows. Parallel and distributed jobs (whether they
are MPI-based, MapReduce-based, etc.) need parallel and distributed storage
systems that keep up with the computation. Furthermore, we highlight the
parallel and distributed computing themes and cast light on them from the
context of our prototype. We also add the Spark discussion proposed by to the
evaluation. The goal is to show how a system such as Cudele would benefit from
more general runtimes. We also move the discussion about the potential for HDFS
and CephFS subtrees in the same namespace and the cost of dynamically changing
consistency and durability semantics on a subtree.

\section*{Cosmetic Fixes}
%Terminology
%- users vs. clients R2
\begin{itemize}
  \item Reviewer 1: fixed user vs. client terminology in Section~{\S}III
  \item Reviewer 1: clarified that the cost of dynamically changing semantics is future work Sections~{\S}III and ~{\S}V.E
  \item Reviewer 1: added source code pointer in Section~{\S}IV
  \item Reviewer 2: remove major takeways in Section~{\S}V
  \item Reviewer 2: remove major cross-references in Section~{\S}V
  \item Reviewer 3: quantified speedups with figure annotations in Section~{\S}V
  \item Reviewers 1 and 3: add servers, network, and storage setups in Section~{\S}V
  \item Reviewer 4: add new section ({\S}V.F) describing how Cudele would work with a system like Spark
\end{itemize}

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,paper}
\end{document}