### 2017-12-22 paper/sections/6_results.tex

\section{Evaluation}
\label{sec:evaluation}

Cudele lets users construct consistency/durability guarantees using
well-established research techniques from other systems; so instead of
evaluating the scalability and performance of the techniques themselves against
other file systems, we show that (1) the mechanisms we propose are useful for
constructing semantics used by real systems, (2) the techniques can work
side-by-side in the same namespace, and (3) the combination of these techniques
can help the system scale further when subtrees are coupled to the correct type
of application.  We scope the evaluation to one metadata server and scale the
number of parallel clients. This type of analysis shows the capacity and
performance of a metadata server with superior metadata protocols, which should
be used to inform load balancing across a metadata cluster. Load balancing
across a cluster of metadata servers with partitioning and replication can be
explored with something like Mantle~\cite{sevilla:sc15-mantle}.

%\begin{table}
%\begin{tabular}{ r | l | l}
%  Cluster  & Hardware               & Software \\\hline
%  In-House & 15 nodes, 4 2GHz CPUs  & Ubuntu 12.04, 1 MON\\
%           & 8GB RAM, SSD           & 1 MDS, 8 OSDs, 8 Clients\\
%  CloudLab & 34 nodes, 16 2.4GHz CPUs & Ubuntu 14.04, 1 MON\\
%           & 128GB RAM, SSD         & 1 MDS, 3 OSDs, 29 Clients
%\end{tabular}
%\caption{In-House used for microbenchmarks, CloudLab for Use Cases. MON is
%monitor daemon, MDS is metadata server daemon, and OSDs are object storage
%daemons.\label{table:clusters}} 
%\end{table}

\newcomment{We run experiments on the same cluster and normalize results to
control for hardware differences and to make our results more generally
applicable. We use a CloudLab cluster of 34 nodes connected with 10Gbit
ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and SSDs.  Each node uses
Ubuntu 14.04 and we develop on Ceph's Jewel release, version 10.2.1, which was
released in May 2016. All daemons run as a single process (the default setting
for Ceph) and the Ceph
components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
object storage daemons, 1 metadata server, and 29 clients.}

\newcomment{We understand concerns about cluster size and network, especially
for systems aimed at HPC but our evaluation focuses on verifying the behavior of
different techniques in the same global namespace. These techniques have
already proven to be effective and scalable in systems that specialize an
entire namespace according to a single optimization strategy (e.g., Lustre,
BatchFS, DeltaFS), so our evaluation explores the behavior of the
consistency/durability mechanisms and their effect with/without isolation.}

\newcomment{To make our results reproducible, this paper adheres to The Popper
Convention~\cite{jimenez_popper_2016}, so experiments presented here are
available in the repository for this article.  Experiments can be examined in
more detail, or even re-run, by visiting the \texttt{[source]} link next to
each figure. That link points to a Jupyter notebook that shows the analysis and
source code for that graph, which points to an experiment and its artifacts.
The source code is available on the Cudele
branch\footnote{\newcomment{https://github.com/michaelsevilla/ceph/tree/cudele}}
of our Ceph fork.} \addressesissue{1}

\subsection{Microbenchmarks}
\label{sec:microbenchmarks}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
\newcomment{Overhead of processing 100K events for each mechanism in
Table~\ref{table:mechanisms}. Results are normalized to the runtime of writing
events to client memory. The bars on the far right show the overhead of
building consistency/durability semantics of real world systems.}
\oldcomment{Performance of each mechanism (left) and building the consistency/durability
semantics of real-world systems (right) for 100K files creates from a single
client. Results are normalized to the runtime of writing events to the client's
in-memory journal.}\label{fig:composable-mechanisms}}
\end{figure}

%\begin{figure}[tb]
%\centering
%\includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
%\caption{ [\href{https://...}{source}] Microbenchmark: storage footprint on
%client. The size of the client's journal scales with the number of
%updates.\label{fig:behavior-journal-size}}
%\end{figure}

Figure~\ref{fig:composable-mechanisms} shows the runtime of the Cudele
mechanisms for a single client creating files in the same directory, normalized
to the time it takes to write 100K file create updates to the client's
in-memory journal ({\it i.e.} the Append Client Journal mechanism).
Stream is an approximation of the overhead and is calculated by subtracting
the runtime of the job with the journal turned off from the runtime with the
journal turned on.  100K is the maximum recommended size of a directory in
CephFS; preliminary experiments with larger directory sizes show memory
problems.

{\it Poorly Scaling Data Structures:} Despite doing the same amount of work,
mechanisms that rely on poorly scaling data structures \oldcomment{have large
slowdowns}\newcomment{can slowdown performance by {\it orders of magnitude}}
for the larger number of creates. For example, RPCs has a \(90\times\) slowdown
because it relies on the internal CephFS directory data structures. It is a
well-known problem that directory data structures do not scale when creating
files in the same directory~\cite{ren:sc2014-indexfs} and any mechanism that
uses these data structures will experience similar slowdowns. Other mechanisms
that write events to a journal) experience a much less drastic slowdown because
the journal data structure does not need to be scanned for every operation.
Events are written to the end of the journal without even checking the validity
({\it e.g.}, if the file already exists for a create), which is another form of
relaxed consistency because the file system assumes the application has
resolved conflicting updates in a different way.

% RPCs vs. apply: calls to metadata server vs. RADOS
{\it Overhead of RPCs:} RPCs is \(66\times\) slower than Volatile
Apply because sending individual metadata updates over the network is costly.
While RPCs sends a request for every file create, Volatile Apply
writes all the updates to the in-memory journal and applies them to the
in-memory data structures in the metadata server. While communicating the
decoupled namespace directly to the metadata server is faster, communicating
through the object store (Nonvolatile Apply) is \(10\times\) slower.

% TODO: why is apply so slow.  
% apply: no CephFS changes, pulls/pushes same RADOS obj.
% v_apply vs. apply/persist: communicating through RADOS
{\it Overhead of Nonvolatile Apply:} The cost of Nonvolatile
Apply is much larger than all the other mechanisms.  That mechanism was not
implemented as part of Cudele -- it was a debugging and recovery tool packaged
with CephFS. It works by iterating over the updates in the journal and pulling
all objects that {\it may} be affected by the update.  This means that two
objects are repeatedly pulled, updated, and pushed: the object that houses the
experiment directory and the object that contains the root directory ({\it
i.e.} \texttt{/}).  The cost of communicating through the object store is shown
by comparing the runtime of Volatile Apply + Global Persist to
Nonvolatile Apply. These two operations end up with the same final metadata
state but using Nonvolatile Apply is clearly inferior.

% persist vs. save: one disk vs. many
{\it Parallelism of the Object Store:} Comparing Local and Global
Persist demonstrates the bandwidth advantages of storing the journal in a
distributed object store. The Global Persist
performance is \(1.5\times\) faster because the object store is leveraging the
collective bandwidth of the disks in the cluster. This benefit comes from the
object store itself but should be acknowledged when making decisions for the
application; the bandwidth of the object store can help mitigate the overheads of
globally persisting metadata updates.

{\it Journal Size:} We measure the amount of storage per journal update to be
about 2.5KB. The storage footprint scales linearly with the number of
metadata creates and suggests that updates for a million files would be 2.5KB
\(*\) 1 million files \(=\) 2.38GB. Transfer times for payloads of this size in
most HPC/data center networks are reasonable.

\oldcomment{\textbf{Takeaway}: measuring the mechanisms individually shows that their
overheads and costs can differ {\it by orders of magnitude}. We also show that
some mechanisms, like Nonvolatile Apply, are not worthwhile as currently
implemented.}

\begin{figure*}[tb]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
      \newcomment{parallel creates on clients}
      }\label{fig:mergescale}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
      \newcomment{block interference}
      }\label{fig:block-allow}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
      \newcomment{syncing to global namespace}
      }\label{fig:slowdown-sync}
  \end{subfigure}
\caption{\oldcomment{ The performance and features of Cudele}\newcomment{Cudele
performance}. (a) \newcomment{shows the speedup of decoupled namespaces over
RPCs; \texttt{create} is the throughput of clients creating files in-parallel
and writing updates locally; \texttt{create+merge} includes the time to merge
updates at the metadata server. }\oldcomment{shows the cost of merging client
journals at the metadata server. Shipping and merging journals of
updates}\newcomment{Decoupled namespaces} scale better than RPCs because there
are less messages and consistency/durability code paths are bypassed. (b) shows
how the allow/block API isolates directories from interfering clients. (c) is
the slowdown of a single client syncing updates to the global namespace. The
inflection point is the trade-off of frequent updates vs. larger journal
files.}
\label{fig:use-cases}
\end{figure*}

\subsection{Use Case 1: Creates in the Same Directory}
\label{sec:use-case-1}

First we show how we can build application-specific subtrees by composing
mechanisms and that \oldcomment{this approach}\newcomment{coupling
well-established techniques to specific applications} results in a more a
scalable global namespace.  We start with clients creating files in private
directories because this workload is heavily studied in
HPC~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs, patil:fast2011-giga,
zheng:pdsw2014-batchfs, sevilla:sc15-mantle}, mostly due to
checkpoint-restart~\cite{bent_plfs_2009}.  But the workload also appears in
cloud workloads~\cite{xiao:socc15-shardfs}, where systems like Hadoop use the
file abstraction to exchange work units to workers or to indicate when jobs
complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
uncompressing an archive ({\it e.g.}, \texttt{tar xzf}), where the file system
services a flash crowd of creates across all directories as shown in
Figure~\ref{fig:overhead-creates}.  The workload is clients creating 100K files
in private directories in the same global namespace.

%\begin{figure}[tb]
%\centering
%\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-weak.png}
%\caption{ [\href{https://...}{source}] Use Case 1: the RPC per metadata update
%(strong consistency) has a large overhead compared to decoupled namespaces
%(weak consistency.)\label{fig:slowdown-strong-weak}}
%\end{figure}

The graph on the right of Figure~\ref{fig:composable-mechanisms} shows how
applications can compose mechanisms together to get the consistency/durability
guarantees they need in a global namespace.  We label the \(x\)-axis with
systems that employ these semantics, as described in
Figure~\ref{fig:subtree-policies}.  Again, the runtime is normalized to
creating files in the client's in-memory journal.  We make no guarantees
during execution of the mechanisms or when transitioning semantics -- the
semantics are guaranteed {\it once the mechanism completes}. So if servers fail
during a mechanism, metadata or data may be lost.

\oldcomment{\textbf{Takeaway}: we confirm the performance benefits of other
well-established research systems but, more importantly, we show that the
Cudele mechanisms we propose are useful for building and evaluating different
consistency/durability semantics.\\}

To show the contention at the metadata server, in Figure~\ref{fig:mergescale}
we scale the number of clients for the RPC per operation strategy and the
decoupled namespace strategy. For RPCs, we increase the number of clients
making requests to the metadata server and for decoupled we increase the number
of concurrent clients operating in parallel. So for the 30 clients data point,
we are measuring the operations per second for 30 client journals that land on
the metadata server at the same time. The ``create + merge" bar adds the time
it takes for clients to generate the journal of events in parallel. The
experiment is run three times but the standard deviation is too small to see.

%Merge requests are serialized at the metadata server; merging concurrently is
%an optimization for future work. The current implementation has race conditions
%because the metadata server recovery code was not designed to run in parallel
%({\it i.e.} the metadata server never recovers with multiple threads). As a
%result, the metadata server versions inodes and directory entries to ensure
%that the recovered metadata is consistent.

This workload scales further with decoupled namespaces because the metadata
server is not overwhelmed with RPC requests. Compared to
Figure~\ref{fig:overhead-c} we can scale to 40 clients which is more than
double the capacity of a single metadata server using the RPC strategy to
maintain strong consistency\footnote{\newcomment{Note that for the 20 clients
bar in Figure~\ref{fig:mergescale}, we plot the throughput for using 18 clients
(the maximum number of clients supported by RPCs)}}. Related work has pointed
to on-disk metadata format as the primary factor for improved scalability; but
our results show that the weakened consistency and durability enables the
scalability improvements. The on-disk metadata formats in this experiment are
the same for all schemes.

The performance of decoupled namespaces is better than RPCs because we place no
restrictions on the validity of metadata inserted into the journal and avoid
touching poorly scaling data structures. We can scale linearly if we weaken our
consistency by never checking for the existence of files before creating files
({\it i.e.} only append \texttt{open()} requests to the journal).  When the
updates are merged by the metadata server into the in-memory metadata store
they never scan growing data structures.  Had we implemented the Append Client
Journal mechanism to include \texttt{lookup()} commands before \texttt{open()}
requests, we would have seen the poor scaling that we see with the RPCs
mechanism.

%% DeltaFS vs. BatchFS vs. POSIX IO: decoupled is faster
%% (5-7x) vs. (90-104x) slower
%% Does not scale?
%% Strong Consistency > 10X slower than Weak Consistency 
%{\it Speedups of Decoupled Namespaces:} Weak consistency uses the
%decoupled namespace strategy and shows up to a \(20\times\) speedup over the
%traditional namespaces that use RPCs. Compared to the baseline the slowdown is
%\(5-7\times\) for Strong Consistency, which emulates BatchFS and
%\(90-104\times\) for Weak Consistency, which emulates DeltaFS.

%{\it Durability \(<<\) Consistency:} The performance of these semantics
%suggests that the overhead of consistency is much larger than the overhead of
%durability. This conclusion should be stronger as we scale the number of files
%because the cost of streaming the journal into the object store is constant.
%
% Weak Consistency benefit not due to metadata format

\oldcomment{\textbf{Takeaway}: the ability to couple well-established techniques to
specific applications allows the global namespace to scale further and perform
better. In our case, RPCs overwhelm the system but decoupling/shipping a
journal of updates improves scalability by 2\(\times\).}

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: weak vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: weak, local
%  - e.g., POSIX IO: strong, global
%  - e.g., PLFS: no consistency
%- durability: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 metadata server, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

\subsection{Use Case 2: Creates with Interfering Client}

Next we show how Cudele can be programmed to block interfering
clients\oldcomment{ using the same problematic workload from
Figure~\ref{fig:overhead-b}. In this workload, c}\newcomment{, which lets
applications control isolation to get better and more reliable performance.
C}lients create files in private directories while another client interferes by
creating \oldcomment{more}\newcomment{1000} files in each directory. This
introduces false sharing and the metadata server revokes capabilities on
directories touched by the interfering client. While HPC tries to avoid these
situations with workflows~\cite{zheng:pdsw2014-batchfs,
zheng:pdsw2015-deltafs}, it still happens in distributed file systems when
users unintentionally access directories in a shared file system.
\oldcomment{We only scale to 15 clients because the performance variability of
a nearly overloaded metadata server, in our case 18 clients, is too high.}

\newcomment{We switch back to clients doing RPCs per operation with journaling
enabled, to mirror the setup from the problem presented in
Figure~\ref{fig:overhead-b}.} Figure~\ref{fig:block-allow} plots the overhead
of the slowest client \oldcomment{and the error bars are the standard deviation
of 3 runs.  ``Interfere (Fig 3c)" is the result from
Figure~\ref{fig:overhead-b}; ``Interfere" uses the allow/block API to return
\texttt{-EBUSY} to interfering clients; and ``Isolated" is the baseline
performance without an interfering.  Each subtree has RPCs and Stream enabled.
Results are normalized to the runtime of a single client creating files in a
directory.  Note that IndexFS has a similar behavior to ``interfere" with
leases, except clients block.  }\newcomment{normalized to 1 client that creates
100K files in isolation (about 513 creates/sec). ``Interference" and ``no
interference" is the performance with and without an interfering client
touching files in every directory, respectively. The goal is to explicitly
isolate clients so that performance is similar to the ``no interfere" curve,
which has lower slowdowns (on average, \(1.42\times\) per client compared to
\(1.67\times\) per client for ``interfere") and less variability (on average, a
standard deviation of 0.06 compared to 0.44 for ``interfere").  ``Block
interference" uses the Cudele API to block interfering clients with
\texttt{-EBUSY}. That curve shows slowdowns similar to ``interfere" for smaller
clusters because the overhead to reject requests is more evident when the
metadata server is underloaded; but the slowdown (\(1.34\times\) per client)
and variability (0.09) look very similar to ``no interfere" for larger
clusters. Clients can block interfering clients to get the same performance as
isolated clients but there is a non-negligible overhead for rejecting requests
when the metadata server is not operating at peak efficiency.}

\oldcomment{We draw three conclusions: (1) clients that use the API to block interfering
clients  get the same performance as isolated clients, (2) there is a
negligible effect on performance for the extra work the metadata server does to
return \texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
negligible effect on performance.

\textbf{Takeaway}: the API lets users isolate directories when applications
need better and more reliable performance. Blocking updates is an effective way of
controlling consistency.}

\subsection{Use Case 3: Read while Writing}

The final use case shows how the API gives users fine-grained control of the
consistency semantics\newcomment{ to support current practices and experimental
workflows}. \oldcomment{The use case is that u}\newcomment{U}sers often
leverage the file system to check the progress of jobs using \texttt{ls} even
though this operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
eshel:fast10-panache} . The number of files or size of the files is indicative
of the progress. This practice is not too different from systems that use the
file system to manage the progress of jobs; Hadoop writes to temporary files,
renames them when complete, and creates a ``DONE" file to indicate to the
runtime that the task did not fail and should not be re-scheduled on another
node. In this scenario, Cudele users will not see the progress of decoupled
namespaces since their updates are not globally visible.  To help users judge
the progress of their jobs, Cudele clients have a ``namespace sync" that sends
batches of updates back to the global namespace at regular intervals.

% Does this implementation need to be described in the implementation?
Figure~\ref{fig:slowdown-sync} shows the performance degradation of a single
client writing \newcomment{1 million} updates to a decoupled namespace and
pausing to send updates to the metadata server. The error bars are the standard
deviations of 3 runs. We scale the namespace sync interval to show the
trade-off of frequently pausing or writing large logs of updates.  We use an
idle core to log the updates and to do the network transfer. The client only
pauses to fork off a background process, which is expensive as the address
space needs to copied. The alternative is to pause the client completely and
write the update to disk but since this implementation is limited by the speed
of the disk, we choose the memory-to-memory copy of the fork approach.

As expected, syncing namespace updates too frequently has the highest overhead
(up to 9\% overhead if done every second). The optimal sync interval for
performance is 10 seconds because larger intervals must write more updates to
disk and network. For the 25 second interval, the client only pauses 3-4 times
but each sync writes about 278 thousand updates at once, which is a 678MB
journal.

\oldcomment{\textbf{Takeaway}: Syncing namespace updates has up to a 9\%
overhead and can be tuned depending on the users preference but, more
importantly, the API gives users fine-grain control of their
consistency/durability to support current practices or experimental workflows.}

%\subsection{Use Case 4: Reading Large Directories} 
%
%% CITEME
%The final use case is reading large directories. At job completion, scientists
%might use \texttt{ls} again to see the results. This causes great strain on the
%file system as paths need to traversed and the entire directory, with all its
%entries, must be transferred back to the client. Here we show how decoupling a
%large namespace and materializing the view in memory on the client is faster
%than doing RPCs for walks of the file system namespace.
%
%FigureX shows how fast a client can materialize decoupled namespaces in memory.
%This takes the on-disk file system metadata format and parses them into events
%that can be manipulated and appended to. Again, we scale the number of up

\subsection{Discussion and Future Work}

\newcomment{We have shown that our API and framework is an effective
abstraction for letting users custom fit subtrees to applications.  It is more
scalable than the current practice of mounting different storage systems in a
global namespace because there is no need to provision dedicated storage
clusters to applications or move data between these systems.  For example, the
results of a Hadoop job do not need to be migrated into a Ceph file system
(CephFS) for other processing; instead the user can change the semantics of the
HDFS subtree into a CephFS subtree.  This may cause metadata/data movement to
make things strongly consistent again but this is superior to moving all data
across file system boundaries. Our prototype enables studies that adjust these
semantics over {\it time and space}, where subtrees can change their semantics
and migrate around the cluster without ever moving the data they reference.}

Cudele and the experiments we present here prompt many avenues for future work.
First is quantifying the benefits of dynamically changing the semantics of a
subtree from stronger to weaker guarantees (or vice versa), as described in the
introduction. Second is embeddable policies, where child subtrees have
specialized features but still maintain the guarantees  of their parent
subtrees. We already show how applications can compose guarantees and control
their performance, knobs that we did not have before, but embeddable policies
are a way to compose the policies themselves.  For example, a RAMDisk subtree
is POSIX IO-compliant but relaxes durability constraints, so it can reside
under a POSIX IO subtree that is strongly consistent. These embeddable policies
should be controlled and enforced by Cudele.  Finally, performance prediction
benefits from our cost quantification of each mechanism. Applications can use
Cudele to microbenchmark their components and software, similar to what we did
in Section~\S\ref{sec:microbenchmarks}. Using those results, they can predict
how much slower their system will be if they adopt stronger consistency or
durability.  This is a form of co-design that takes a ``dirty-slate" approach
but building just the guarantees the application needs from existing
implementations.  This can also be a verification tool where performance that
varies wildly from the predicted performance can be a red flag that something
is wrong or that the bottleneck is not in the consistency or durability plane.

%Second is the opportunity for adding hardware in strategic places to speed up
%the mechanisms that the application needs. In Figure~\ref{fig:mergescale} the
%performance of merging updates at the metadata server is limited by the speed
%of the disk. So applications that rely on merging many updates should consider
%equipping the metadata server with SSDs or more memory. 
%