### 2017-12-23 paper/sections/6_results.tex

\section{Evaluation}
\label{sec:evaluation}

Cudele lets users construct consistency/durability guarantees using
well-established research techniques from other systems; so instead of
evaluating the scalability and performance of the techniques themselves against
other file systems, we show that (1) the mechanisms we propose are useful for
constructing semantics used by real systems, (2) the techniques can work
side-by-side in the same namespace, and (3) the combination of these techniques
can help the system scale further when subtrees are coupled to the correct type
of application.  \newcomment{These techniques have already proven to be
effective and scalable in systems that specialize an entire namespace according
to a single optimization strategy (e.g., Lustre, BatchFS, DeltaFS), so our
evaluation explores the behavior of the consistency/durability mechanisms and
their effect with/without isolation.}

\newcomment{\indent We run experiments on the same cluster and graph the
standard deviations for three runs (sometimes error bars are too small to see).
We also normalize results to control for hardware differences and to make our
results more generally applicable. We use a CloudLab cluster of 34 nodes
connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and SSDs.
Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel release, version
10.2.1, which was released in May 2016. All daemons run as a single process
(the default setting for Ceph) and the Ceph
components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
object storage daemons, 1 metadata server, and 29 clients.} We scope the
evaluation to one metadata server and scale the number of parallel
clients\newcomment{ each doing 100K operations}\footnote{This is the maximum
recommended size of a directory in CephFS; preliminary experiments with larger
directory sizes show memory problems.}. This type of analysis shows the
capacity and performance of a metadata server with superior metadata protocols,
which should be used to inform load balancing across a metadata cluster. Load
balancing across a cluster of metadata servers with partitioning and
replication can be explored with something like
Mantle~\cite{sevilla:sc15-mantle}.

\newcomment{\indent To make our results reproducible, this paper adheres to The
Popper Convention~\cite{jimenez_popper_2016}, so experiments presented here are
available in the repository for this article.  Experiments can be examined in
more detail, or even re-run, by visiting the \texttt{[source]} link next to
each figure. That link points to a Jupyter notebook that shows the analysis and
source code for that graph, which points to an experiment and its artifacts.
The source code is available on the Cudele
branch\footnote{\newcomment{https://github.com/michaelsevilla/ceph/tree/cudele}}
of our Ceph fork.} \addressesissue{1}

%\newcomment{\indent We understand concerns about cluster size and network, especially
%for systems aimed at HPC but our evaluation focuses on verifying the behavior
%of different techniques in the same global namespace.

%\begin{table}
%\begin{tabular}{ r | l | l}
%  Cluster  & Hardware               & Software \\\hline
%  In-House & 15 nodes, 4 2GHz CPUs  & Ubuntu 12.04, 1 MON\\
%           & 8GB RAM, SSD           & 1 MDS, 8 OSDs, 8 Clients\\
%  CloudLab & 34 nodes, 16 2.4GHz CPUs & Ubuntu 14.04, 1 MON\\
%           & 128GB RAM, SSD         & 1 MDS, 3 OSDs, 29 Clients
%\end{tabular}
%\caption{In-House used for microbenchmarks, CloudLab for Use Cases. MON is
%monitor daemon, MDS is metadata server daemon, and OSDs are object storage
%daemons.\label{table:clusters}} 
%\end{table}

\subsection{Microbenchmarks}
\label{sec:microbenchmarks}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
\newcomment{Overhead of processing 100K create events for each mechanism in
Table~\ref{table:mechanisms}. Results are normalized to the runtime of writing
events to client memory. The bars on the far right show the overhead of
building consistency/durability semantics of real world systems.}
\oldcomment{Performance of each mechanism (left) and building the consistency/durability
semantics of real-world systems (right) for 100K files creates from a single
client. Results are normalized to the runtime of writing events to the client's
in-memory journal.}\label{fig:composable-mechanisms}}
\end{figure}

%\begin{figure}[tb]
%\centering
%\includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
%\caption{ [\href{https://...}{source}] Microbenchmark: storage footprint on
%client. The size of the client's journal scales with the number of
%updates.\label{fig:behavior-journal-size}}
%\end{figure}

\oldcomment{Figure~\ref{fig:composable-mechanisms} shows the runtime of the
Cudele mechanisms for a single client creating files in the same directory, }
\newcomment{We measure the overhead of each mechanism by having 1 client create
100K files in a directory for various subtree configurations.
Figure~\ref{fig:composable-mechanisms} shows the time that it takes each Cudele
mechanism to process all metadata events.  Results are} normalized to the time
it takes to write 100K file create updates to the client's in-memory journal
({\it i.e.} the Append Client Journal mechanism).  \newcomment{The first graph
groups the consistency mechanisms, the second groups the durability mechanisms,
and the third has compositions representing real-world systems.} Stream is an
approximation of the overhead and is calculated by subtracting the runtime of
the job with the journal turned off from the runtime with the journal turned
on.  \oldcomment{100K is the maximum recommended size of a directory in CephFS;
preliminary experiments with larger directory sizes show memory problems.}

{\it Poorly Scaling Data Structures:} Despite doing the same amount of work,
mechanisms that rely on poorly scaling data structures have large
slowdowns\oldcomment{ for the larger number of creates}. For example, RPCs has
a \oldcomment{\(90\times\)}\newcomment{\(18\times\)} slowdown because this
technique relies on the internal directory data structures. It is a well-known
problem that directory data structures do not scale when creating files in the
same directory~\cite{ren:sc2014-indexfs} and any mechanism that uses these data
structures will experience similar slowdowns. Other mechanisms that write
events to a journal experience a much less drastic slowdown because the journal
data structure does not need to be scanned for every operation.  Events are
written to the end of the journal without even checking the validity ({\it
e.g.}, if the file already exists for a create), which is another form of
relaxed consistency because the file system assumes the application has
resolved conflicting updates in a different way.

% RPCs vs. apply: calls to metadata server vs. RADOS
{\it Overhead of RPCs:} RPCs is
\oldcomment{\(66\times\)}\newcomment{\(19.9\times\)} slower than Volatile Apply
because sending individual metadata updates over the network is costly.  While
RPCs sends a request for every file create, Volatile Apply \oldcomment{writes
all the updates to the in-memory journal and applies them}\newcomment{writes
directly} to the in-memory data structures in the metadata server. While
communicating the decoupled namespace directly to the metadata server
\newcomment{with Volatile Apply} is faster, communicating through the object
store \newcomment{with Nonvolatile Apply} is
\oldcomment{\(10\times\)}\newcomment{\(78\times\)} slower.

% TODO: why is apply so slow.  apply: no CephFS changes, pulls/pushes same
% RADOS obj.  v_apply vs. apply/persist: communicating through RADOS
{\it Overhead of Nonvolatile Apply:} The cost of Nonvolatile Apply is much
larger than all the other mechanisms.  That mechanism was not implemented as
part of Cudele -- it was a debugging and recovery tool packaged with CephFS. It
works by iterating over the updates in the journal and pulling all objects that
{\it may} be affected by the update.  This means that two objects are
repeatedly pulled, updated, and pushed: the object that houses the experiment
directory and the object that contains the root directory ({\it i.e.},
\texttt{/}).  The cost of communicating through the object store is shown by
comparing the runtime of Volatile Apply + Global Persist
\newcomment{(\(1.3\times\))} to Nonvolatile Apply\newcomment{ (\(78\times\))}.
These two operations end up with the same final metadata state but using
Nonvolatile Apply is clearly inferior.

% persist vs. save: one disk vs. many
{\it Parallelism of the Object Store:} \newcomment{Stream has the highest
slowdown at \(2.4\times\) because the overhead of maintaining and streaming the
journal is incurred by the metadata server}.  Comparing Local and Global
Persist demonstrates the bandwidth advantages of storing the journal in a
distributed object store. The Global Persist performance is \newcomment{only
\(0.2\times\) slower than Local Persist because Global
Persist}\oldcomment{\(1.5\times\) faster because the object store} is
leveraging the collective bandwidth of the disks in the cluster. This benefit
comes from the object store itself but should be acknowledged when making
decisions for the application; the bandwidth of the object store can help
mitigate the overheads of globally persisting metadata updates. The storage per
journal update is about 2.5KB. So the storage footprint scales linearly with
the number of metadata creates and suggests that updates for a million files
would be \newcomment{2.38GB}\oldcomment{2.5KB \(*\) 1 million files \(=\)
2.38GB. Transfer times for payloads of this size in most HPC/data center
networks are reasonable.}

\oldcomment{\textbf{Takeaway}: measuring the mechanisms individually shows that their
overheads and costs can differ {\it by orders of magnitude}. We also show that
some mechanisms, like Nonvolatile Apply, are not worthwhile as currently
implemented.}

\begin{figure*}[tb]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
      \newcomment{parallel creates on clients}
      }\label{fig:mergescale}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
      \newcomment{block interference}
      }\label{fig:block-allow}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
      \caption{
      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
      \newcomment{syncing to global namespace}
      }\label{fig:slowdown-sync}
  \end{subfigure}
\caption{\oldcomment{ The performance and features of Cudele}\newcomment{Cudele
performance}. (a) \newcomment{shows the speedup of decoupled namespaces over
RPCs; \texttt{create} is the throughput of clients creating files in-parallel
and writing updates locally; \texttt{create+merge} includes the time to merge
updates at the metadata server. }\oldcomment{shows the cost of merging client
journals at the metadata server. Shipping and merging journals of
updates}\newcomment{Decoupled namespaces} scale better than RPCs because there
are less messages and consistency/durability code paths are bypassed. (b) shows
how the allow/block API isolates directories from interfering clients. (c) is
the slowdown of a single client syncing updates to the global namespace. The
inflection point is the trade-off of frequent updates vs. larger journal
files.}
\label{fig:use-cases}
\end{figure*}

\newcomment{{\it Composing Mechanisms}: The graph on the right of
Figure~\ref{fig:composable-mechanisms} shows how applications can compose
mechanisms together to get the consistency/durability guarantees they need in a
global namespace.  We label the \(x\)-axis with systems that employ these
semantics, as described in Figure~\ref{fig:subtree-policies}.  Again, the
runtime is normalized to creating files in the client's in-memory journal.  We
make no guarantees during execution of the mechanisms or when transitioning
semantics -- the semantics are guaranteed {\it once the mechanism completes}.
So if servers fail during a mechanism, metadata or data may be lost. This graph
shows how we can build application-specific subtrees by composing mechanisms
and that coupling well-established techniques to specific applications results
in a more scalable global namespace.}

\subsection{Use Case 1: Creates in the Same Directory}
\label{sec:use-case-1}

\oldcomment{Next we show how we can build application-specific subtrees by composing
mechanisms and that \oldcomment{this approach}\newcomment{coupling
well-established techniques to specific applications} results in a more a
scalable global namespace.  }

We start with clients creating files in private directories because this
workload is heavily studied in HPC~\cite{weil:sc2004-dyn-metadata,
ren:sc2014-indexfs, patil:fast2011-giga, zheng:pdsw2014-batchfs,
sevilla:sc15-mantle}, mostly due to checkpoint-restart~\cite{bent_plfs_2009}.
But the workload also appears in cloud workloads~\cite{xiao:socc15-shardfs},
where systems like Hadoop use the file abstraction to exchange work units to
workers or to indicate when jobs
complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
uncompressing an archive ({\it e.g.}, \texttt{tar xzf}), where the file system
services a flash crowd of creates across all directories as shown in
Figure~\ref{fig:overhead-creates}. \oldcomment{ The workload is clients creating 100K files
in private directories in the same global namespace. }

%\begin{figure}[tb]
%\centering
%\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-weak.png}
%\caption{ [\href{https://...}{source}] Use Case 1: the RPC per metadata update
%(strong consistency) has a large overhead compared to decoupled namespaces
%(weak consistency.)\label{fig:slowdown-strong-weak}}
%\end{figure}

\oldcomment{The graph on the right of Figure~\ref{fig:composable-mechanisms} shows how
applications can compose mechanisms together to get the consistency/durability
guarantees they need in a global namespace.  We label the \(x\)-axis with
systems that employ these semantics, as described in
Figure~\ref{fig:subtree-policies}.  Again, the runtime is normalized to
creating files in the client's in-memory journal.  We make no guarantees
during execution of the mechanisms or when transitioning semantics -- the
semantics are guaranteed {\it once the mechanism completes}. So if servers fail
during a mechanism, metadata or data may be lost.}

\oldcomment{\textbf{Takeaway}: we confirm the performance benefits of other
well-established research systems but, more importantly, we show that the
Cudele mechanisms we propose are useful for building and evaluating different
consistency/durability semantics.\\}

\oldcomment{To show the contention at the metadata server, }In
Figure~\ref{fig:mergescale} we scale the number of clients \newcomment{each
doing 100K file creates in their own directories for three types of subtrees:
one with strong consistency and global durability (RPCs), one with invisible
consistency and local durability (decoupled: create), and one with weak
consistency and local durability (decoupled: create + merge). Results are
normalized to 1 client that creates 100K files using RPCs (about 549
creates/sec). As opposed to earlier graphs in
Section~\S\ref{sec:posix-overheads} that plotted the throughput of the slowest
client, Figure~\ref{fig:mergescale} plots the throughput from the perspective
of the metadata server. Plotting this way is easier to understand because of
how we normalize but the speedups over the RPC approach are the same, whether
we look at the slowest client or not.}

\newcomment{When the metadata server is operating at peak efficiency between 5
and 20 clients, ``decoupled: merge + create" outperforms ``RPCs" by
\(3.37\times\) because ``decoupled: merge + create" uses a relaxed form of
consistency and leverages bulk updates just liked
DeltaFS~\cite{zheng:pdsw2015-deltafs}.  The ``RPCs" and ``decoupled: merge +
create" curves flatten out at a slowdown of \(4.5\times\) and \(15\times\),
respectively, because more clients increases the load at the metadata server.
Alternatively, the ``decoupled: create" curve scales linearly with the number
of concurrent clients because clients operate in parallel and write updates
locally. At 20 clients, we observe a \(91.7\times\) speedup for ``decoupled:
create" over RPCs.}\oldcomment{ for the RPC per operation strategy and the
decoupled namespace strategy.  For RPCs, scaling the number of clients
increases the load on the metadata server and for decoupled we increase the
number of concurrent clients operating in parallel.  Related work has pointed
to on-disk metadata format as the primary factor for improved scalability; but
our results show that the weakened consistency and durability enables the
scalability improvements. The on-disk metadata formats in this experiment are
the same for all schemes.}

\newcomment{Another reason that} the performance of decoupled namespaces is
better than RPCs \newcomment{is because decoupled namespaces (1)} place no
restrictions on the validity of metadata inserted into the journal and
\newcomment{(2)} avoid touching poorly scaling data structures. We can scale
linearly if we weaken our consistency by never checking for the existence of
files before creating files ({\it i.e.} only append \texttt{open()} requests to
the journal).  When the updates are merged by the metadata server into the
in-memory metadata store they never scan growing data structures.  Had we
implemented the \oldcomment{Append Client Journal mechanism}\newcomment{client}
to include \texttt{lookup()} commands before \texttt{open()} requests, we would
have seen the poor scaling \newcomment{behavior} that we see with the RPCs
mechanism.

\newcomment{One final point about the ``decoupled: merge + create" curve is
that the results in the graph are pessimistic. They model a scenario in which
all client journals arrive at the same time.} So for the 20 clients data point,
we are measuring the operations per second for 20 client journals that land on
the metadata server at the same time. \newcomment{Had we added infrastructure
to overlay journal arrivals or time client sync intervales, we could have
scaled more closely to ``decoupled: create".}\oldcomment{The ``create + merge"
bar adds the time it takes for clients to generate the journal of events in
parallel. ALSO DISCUSS 3.7X!!}

%Merge requests are serialized at the metadata server; merging concurrently is
%an optimization for future work. The current implementation has race conditions
%because the metadata server recovery code was not designed to run in parallel
%({\it i.e.} the metadata server never recovers with multiple threads). As a
%result, the metadata server versions inodes and directory entries to ensure
%that the recovered metadata is consistent.

\oldcomment{This workload scales further with decoupled namespaces because the metadata
server is not overwhelmed with RPC requests. Compared to
Figure~\ref{fig:overhead-c} we can scale to 40 clients which is more than
double the capacity of a single metadata server using the RPC strategy to
maintain strong consistency Note that for the 20 clients
bar in Figure~\ref{fig:mergescale}, we plot the throughput for using 18 clients
(the maximum number of clients supported by RPCs). }

\oldcomment{\textbf{Takeaway}: the ability to couple well-established techniques to
specific applications allows the global namespace to scale further and perform
better. In our case, RPCs overwhelm the system but decoupling/shipping a
journal of updates improves scalability by 2\(\times\).}

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: weak vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: weak, local
%  - e.g., POSIX IO: strong, global
%  - e.g., PLFS: no consistency
%- durability: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 metadata server, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

\subsection{Use Case 2: Creates with Interfering Client}

Next we show how Cudele can be programmed to block interfering
clients\oldcomment{ using the same problematic workload from
Figure~\ref{fig:overhead-b}. In this workload, c}\newcomment{, which lets
applications control isolation to get better and more reliable performance.
C}lients create 100K files in their own directories while another client interferes by
creating \oldcomment{more}\newcomment{1000} files in each directory. This
introduces false sharing and the metadata server revokes capabilities on
directories touched by the interfering client. While HPC tries to avoid these
situations with workflows~\cite{zheng:pdsw2014-batchfs,
zheng:pdsw2015-deltafs}, it still happens in distributed file systems when
users unintentionally access directories in a shared file system.
\oldcomment{We only scale to 15 clients because the performance variability of
a nearly overloaded metadata server, in our case 18 clients, is too high.}

\newcomment{We switch back to clients doing RPCs per operation with journaling
enabled, to mirror the setup from the problem presented in
Figure~\ref{fig:overhead-b}.} Figure~\ref{fig:block-allow} plots the overhead
of the slowest client \oldcomment{and the error bars are the standard deviation
of 3 runs.  ``Interfere (Fig 3c)" is the result from
Figure~\ref{fig:overhead-b}; ``Interfere" uses the allow/block API to return
\texttt{-EBUSY} to interfering clients; and ``Isolated" is the baseline
performance without an interfering.  Each subtree has RPCs and Stream enabled.
Results are normalized to the runtime of a single client creating files in a
directory.  Note that IndexFS has a similar behavior to ``interfere" with
leases, except clients block.  }\newcomment{normalized to 1 client that creates
100K files in isolation (about 513 creates/sec). ``Interference" and ``no
interference" is the performance with and without an interfering client
touching files in every directory, respectively. The goal is to explicitly
isolate clients so that performance is similar to the ``no interfere" curve,
which has lower slowdowns (on average, \(1.42\times\) per client compared to
\(1.67\times\) per client for ``interfere") and less variability (on average, a
standard deviation of 0.06 compared to 0.44 for ``interfere").  ``Block
interference" uses the Cudele API to block interfering clients with
\texttt{-EBUSY}. That curve shows slowdowns similar to ``interfere" for smaller
clusters because the overhead to reject requests is more evident when the
metadata server is underloaded; but the slowdown (\(1.34\times\) per client)
and variability (0.09) look very similar to ``no interfere" for larger
clusters. Clients can block interfering clients to get the same performance as
isolated clients but there is a non-negligible overhead for rejecting requests
when the metadata server is not operating at peak efficiency.}

\oldcomment{We draw three conclusions: (1) clients that use the API to block interfering
clients  get the same performance as isolated clients, (2) there is a
negligible effect on performance for the extra work the metadata server does to
return \texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
negligible effect on performance.}

\oldcomment{\textbf{Takeaway}: the API lets users isolate directories when applications
need better and more reliable performance. Blocking updates is an effective way of
controlling consistency.}

\subsection{Use Case 3: Read while Writing}

The final use case shows how the API gives users fine-grained control of the
consistency semantics \newcomment{to support current practices and experimental
workflows}. \oldcomment{The use case is that u}\newcomment{U}sers often
leverage the file system to check the progress of jobs using \texttt{ls} even
though this operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
eshel:fast10-panache}. The number of files or size of the files is indicative
of the progress. This practice is not too different from systems that use the
file system to manage the progress of jobs; Hadoop writes to temporary files,
renames them when complete, and creates a ``DONE" file to indicate to the
runtime that the task did not fail and should not be re-scheduled on another
node. In this scenario, Cudele users will not see the progress of decoupled
namespaces since their updates are not globally visible.  To help users judge
the progress of their jobs, Cudele clients have a ``namespace sync" that sends
batches of updates back to the global namespace at regular intervals.

% Does this implementation need to be described in the implementation?
\newcomment{We configure a subtree as a decoupled namespace with invisible
consistency, local durability, and partial updates enabled.}
Figure~\ref{fig:slowdown-sync} shows the performance degradation of a single
client writing \newcomment{1 million} updates to the decoupled namespace and
pausing to send updates to the metadata server. \oldcomment{The error bars are
the standard deviations of 3 runs.} We scale the namespace sync interval to
show the trade-off of frequently pausing or writing large logs of updates.  We
use an idle core to log the updates and to do the network transfer. The client
only pauses to fork off a background process, which is expensive as the address
space needs to be copied. The alternative is to pause the client completely and
write the update to disk but since this implementation is limited by the speed
of the disk, we choose the memory-to-memory copy of the fork approach.

As expected, syncing namespace updates too frequently has the highest overhead
(up to 9\% overhead if done every second). The optimal sync interval for
performance is 10 seconds, which only incurs 2\% overhead), because larger
intervals must write more updates to disk and network. For the 25 second
interval, the client only pauses 3-4 times but each sync writes about 278
thousand updates at once, which is a 678MB journal.

\oldcomment{\textbf{Takeaway}: Syncing namespace updates has up to a 9\%
overhead and can be tuned depending on the users preference but, more
importantly, the API gives users fine-grain control of their
consistency/durability to support current practices or experimental workflows.}

%\subsection{Use Case 4: Reading Large Directories} 
%
%% CITEME
%The final use case is reading large directories. At job completion, scientists
%might use \texttt{ls} again to see the results. This causes great strain on the
%file system as paths need to traversed and the entire directory, with all its
%entries, must be transferred back to the client. Here we show how decoupling a
%large namespace and materializing the view in memory on the client is faster
%than doing RPCs for walks of the file system namespace.
%
%FigureX shows how fast a client can materialize decoupled namespaces in memory.
%This takes the on-disk file system metadata format and parses them into events
%that can be manipulated and appended to. Again, we scale the number of up

\subsection{Discussion and Future Work}

\newcomment{We have shown that our API and framework is an effective
abstraction for letting users custom fit subtrees to applications.  It is more
scalable than the current practice of mounting different storage systems in a
global namespace because there is no need to provision dedicated storage
clusters to applications or move data between these systems.  For example, the
results of a Hadoop job do not need to be migrated into a Ceph file system
(CephFS) for other processing; instead the user can change the semantics of the
HDFS subtree into a CephFS subtree.  This may cause metadata/data movement to
make things strongly consistent again but this is superior to moving all data
across file system boundaries. Our prototype enables studies that adjust these
semantics over {\it time and space}, where subtrees can change their semantics
and migrate around the cluster without ever moving the data they reference.}

Cudele and the experiments we present here prompt many avenues for future work.
First is quantifying the benefits of dynamically changing the semantics of a
subtree from stronger to weaker guarantees (or vice versa), as described in the
introduction. Second is embeddable policies, where child subtrees have
specialized features but still maintain the guarantees  of their parent
subtrees. We already show how applications can compose guarantees and control
their performance, knobs that we did not have before, but embeddable policies
are a way to compose the policies themselves.  For example, a RAMDisk subtree
is POSIX IO-compliant but relaxes durability constraints, so it can reside
under a POSIX IO subtree that is strongly consistent. These embeddable policies
should be controlled and enforced by Cudele.  Finally, performance prediction
benefits from our cost quantification of each mechanism. Applications can use
Cudele to microbenchmark their components and software, similar to what we did
in Section~\S\ref{sec:microbenchmarks}. Using those results, they can predict
how much slower their system will be if they adopt stronger consistency or
durability.  This is a form of co-design that takes a ``dirty-slate" approach
but building just the guarantees the application needs from existing
implementations.  This can also be a verification tool where performance that
varies wildly from the predicted performance can be a red flag that something
is wrong or that the bottleneck is not in the consistency or durability plane.

%Second is the opportunity for adding hardware in strategic places to speed up
%the mechanisms that the application needs. In Figure~\ref{fig:mergescale} the
%performance of merging updates at the metadata server is limited by the speed
%of the disk. So applications that rely on merging many updates should consider
%equipping the metadata server with SSDs or more memory. 
%