### 2018-01-09 paper/sections/1_introduction.tex

\section{Introduction}

% What is the problem?
File system metadata services in HPC and large-scale data centers have
scalability problems because common tasks, like
checkpointing~\cite{bent_plfs_2009} or scanning the file
system~\cite{zheng:pdsw2014-batchfs}, contend for the same directories and
inodes. Applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
metadata server for every \newcomment{client}\footnote{In this paper, ``client"
is a storage client or application that interacts with the metadata server,
``administrator" is a system administrator that configures the storage, and
``end-users" interact with the file system via home directories or runtimes.  }
is unreasonable. This problem is exacerbated by current hardware and software
trends: parallel runtimes like Hadoop and Spark use storage systems to
implement important scheduling and timing semantics and  \oldcomment{; for
example, } HPC architectures are transitioning from complex storage stacks with
burst buffer, file system, object store, and tape tiers to more simplified
stacks with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}. These types of trends put pressure on
data access because more requests from different nodes end up hitting the same
layers in parallel\oldcomment{and latencies cannot be hidden while data
migrates across tiers}.

% What is HPC doing?
To address this, developers are relaxing the consistency and durability
semantics in the file system because weaker guarantees are sufficient for their
applications. For example, many HPC batch jobs do not need the strong
consistency that the file system provides, so
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
do more client-side processing and merge updates when the job is done.
Developers in these domains are turning to these non-POSIX IO solutions because
their applications are well-understood ({\it e.g.}, well-defined read/write
phases, synchronization only needed during certain phases, workflows describing
computation, etc.) and because these applications wreak havoc on file systems
designed for general-purpose workloads ({\it e.g.}, checkpoint-restart's N:N
and N:1 create patterns~\cite{bent_plfs_2009}).

\begin{figure}[tb] \centering
\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
\caption{Illustration of subtrees with different semantics co-existing in a
global namespace.  For performance, clients \oldcomment{can} relax consistency/durability on their
subtree ({\it e.g.}, HDFS) or decouple the subtree and move it locally ({\it e.g.}, BatchFS, RAMDisk).
}\label{fig:subtree-policies} \end{figure}

% One example
One popular approach for relaxing consistency and durability is to ``decouple
the namespace", where clients lock the subtree they want exclusive access to as
a way to tell the file system that the subtree is important or may cause
resource contention in the near-future~\cite{grider:pdsw2015-marfs,
zheng:pdsw2015-deltafs, zheng:pdsw2014-batchfs, ren:sc2014-indexfs,
bent:slides-twotiers}. Then the file system can change its internal structure
to optimize performance. For example, the file system could enter a mode where
clients with decoupled directories perform operations locally and bulk merge their
updates at completion. This delayed merge ({\it i.e.} a form of eventual
consistency) and relaxed durability improves performance and scalability by
avoiding the costs of remote procedure calls (RPCs), synchronization, false
sharing, and serialization.  While the performance benefits of decoupling the
namespace are obvious, applications that rely on the file system's guarantees
must be deployed on an entirely different system or re-written to coordinate
strong consistency/durability themselves.

% What did we do
%\footnote{HDFS itself is not directly evaluated in this paper, although the
%semantics and their performance is explored in~\S\ref{sec:use-case-1}.} 
To address this problem, we present an API and framework that lets developers
dynamically control the consistency and durability guarantees for subtrees in
the file system namespace.  Figure~\ref{fig:subtree-policies} shows a potential
setup in our proposed system where a single global namespace has subtrees for
applications optimized with techniques from different state-of-the-art
architectures.  The BatchFS and RAMDisk subtrees are decoupled from the global
namespace and have similar consistency/durability behavior to those systems;
the HDFS subtree has weaker than strong consistency because it lets clients
read files opened for writing~\cite{hakimzadeh:dais14-hdfs-consistency}, which
means that not all updates are immediately seen by all clients; and the POSIX
IO subtree retains the rigidity of POSIX IO's strong consistency.  Subtrees
without policies inherit the consistency/durability semantics of the parent and
future work will examine embeddable or inheritable policies.

Our prototype system, Cudele, achieves this by exposing ``mechanisms" that
\oldcomment{users}\newcomment{administrators} combine to specify their
preferred semantics.  Cudele supports 3 forms of consistency (invisible, weak,
and strong) and 3 degrees of durability (none, local, and global) giving the
\oldcomment{user}\newcomment{administrator} a wide range of policies and
optimizations that can be custom fit to an application. We make the following
contributions:

\begin{enumerate}

  \item A framework/API for assigning consistency/durability policies to
  subtrees in a global namespace; this lets
  \oldcomment{users}\newcomment{administrators} navigate trade-offs of
  different metadata protocols on the same storage system.

  \item

  \oldcomment{This framework lets subtrees with different semantics co-exist in
a global namespace. We show how this} \newcomment{We show that letting
different semantics co-exist in a global namespace scales further and performs
better than systems that use one strategy.}

  \item A prototype that lets \oldcomment{users}\newcomment{administrators}
  custom fit subtrees to applications dynamically.

\end{enumerate}

\oldcomment{The last contribution lays the groundwork for future work on our
prototype. It is more scalable than the current practice of mounting different
storage systems in a global namespace because there is no need to provision
dedicated storage clusters to applications or move data between these systems.
For example, the results of a Hadoop job do not need to be migrated into a Ceph
file system (CephFS) for other processing; instead the
\oldcomment{users}\newcomment{administrator} can change the semantics of the
HDFS subtree into a CephFS subtree.  This may cause metadata/data movement to
make things strongly consistent again but this is superior to moving all data
across file system boundaries. Our prototype enables studies that adjust these
semantics over {\it time and space}, where subtrees can change their semantics
and migrate around the cluster without ever moving the data they reference.}

% Results
The results in this paper confirm the assertions of ``clean-slate" research of
decoupled namespaces; specifically that these techniques drastically improve
performance.  We go a step further by quantifying the costs of traditional file
system approaches to maintaining consistency (3.37\(\times\) slowdown) and
durability (\(2.4\times\) slowdown).  In our prototype, \newcomment{we also
show the benefits of assigning subtree semantics to certain applications such
as checkpoint-restart (\(91.7\times\) speedup if consistency is fully relaxed),
user home directories (within a 0.03 standard deviation from optimal), and
end-users checking for partial results (only a 2\% overhead).} \oldcomment{we
get an 8\(\times\) speedup and can scale to twice as many clients when we
assign a more relaxed form of consistency and durability to a subtree with a
create-heavy workload.}  We use Ceph as a prototyping platform because it is
used in cloud-based and data center systems and has a presence in
HPC~\cite{wang:pdsw13-ceph-hpc}.

%In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the
%cost of POSIX IO consistency and system-defined durability and
%Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
%prototype and API. Section~\ref{sec:implementation} describes the Cudele
%mechanisms and shows how re-using internal subsystems results in an
%implementation of less than 500 lines of code. The evaluation in
%Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
%explored and previously unexplored metadata designs.
%Section~\ref{sec:related-work} places Cudele in the context of other related
%work.